Testing and Isolation Efficacy; Insights from a Simple Epidemic Model

### 2020-12-30 SIR_manuscript.tex
working on intro

One of the main challenges in understanding COVID-19 dynamics is to unravel information available from data. The daily case count is a highly accessible form of data where observations are driven by two clusters of processes: (i) epidemic processes and (ii) testing processes. The challenge is then to separate these processes and understand how they interact. Simple mathematical models have a long history in providing useful insights facing with real epidemiological challenges \citep{ross1911prevention} [maybe add influenza work, ideas?]. We developed a model that incorporates mechanistic processes of epidemic processes and testing. We used the model to study the potential effect of testing strategies on disease dynamics.

\jd{There is a gap here between unravelling data and what we actually do.}
\bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well). I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}

>>>>

The observed dynamics of the COVID-19 epidemic is dirven by the integrated processes including both epidemiological processes (infection, recovery, etc.) and testing processes. In addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine. Specifically, individuals who are in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate. We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the control of the epidemic.

Spereating the epidemiological processes from testing processes through a mathematical modeling excercise provides a tool to investigate the effect of the testing on the epidemic dynamics. Although, this effect is intuitively through isolation and quarantine practices, the sensitivity of the epidemic dynamics to these practices needs to be studied further. During COVID-19 epidemic, the isolation has applied to people who have been tested. Thus, the epidemic dynamics will depend on testing strategies which determines how tests are distributed across epidemiological compartments, or for a given testing intensity how many people in each compartment get tested. Specifically, in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be distributed equally across all compartments.

that the epi effects will depend on \emph{who gets tested}, i.e. how tests are distributed across compartments, or for a given testing intensity how many people in each epidemiological compartment get tested. Then you can go through some examples, i.e. start by saying that in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments \ldots}

% One of the main challenges in understanding COVID-19 dynamics is to unravel information available from data. The daily case count is a highly accessible form of data where observations are driven by two clusters of processes: (i) epidemic processes and (ii) testing processes. The challenge is then to separate these processes and understand how they interact. Simple mathematical models have a long history in providing useful insights facing with real epidemiological challenges \citep{ross1911prevention} [maybe add influenza work, ideas?]. We developed a model that incorporates mechanistic processes of epidemic processes and testing. We used the model to study the potential effect of testing strategies on disease dynamics.

% \jd{There is a gap here between unravelling data and what we actually do.}
% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well). I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}
